---
title: "Main"
author: "Abhijeet Mohite"
date: "09/06/2020"
output: html_document
---

# this is the file where the code will be updated

#this code is used for reading the data from the csv file activiy.csv

```{r,echo=TRUE}
data <- read.csv("activity.csv")
```
#we need to change the date format into date hence we write the code below
```{r,echo=TRUE}
data$date<-as.Date(data$date)
```
#next we need to creaate a histogram of the total number of steps taken each day
```{r, echo=TRUE}

sum_steps<-aggregate(data$steps,by=list(data$date),FUN=sum,na.rm=TRUE)
```

```{r histogram1, echo=TRUE, fig.height=4}
hist(sum_steps$x, 
     breaks=seq(from=0, to=25000, by=2500),
     col="red", 
     xlab="Total number of steps", 
     ylim=c(0, 20), 
     main="Histogram of the total number of steps taken each day\n(NA removed)")
```

#the next is we will calculate the mean and median steps taken each day
```{r,echo=TRUE}
mean(sum_steps$x)

median(sum_steps$x)
```
#next we will create a time series plot of the average number of steps taken
```{r,echo=TRUE}
avg_steps<-aggregate(data$steps,by=list(data$interval),FUN=mean,na.rm=TRUE)

colnames(avg_steps)<-c("interval","steps")

library(ggplot2)
```

```{r timeseries plot, echo=TRUE,fig.height=4}

ggplot(aes(x=interval,y=steps),data=avg_steps)+geom_line()
```
#now we calculate the five minute intervals that on average that every step contains
```{r,echo=TRUE}
avg_steps[avg_steps$steps==max(avg_steps$steps),1]
```
#next we describe the steps for inputting the data
#first we calculate and report the number of missing values in the data

```{r,echo=TRUE}
sum(is.na(data$steps))
```
#replace NA with the means of the steps
```{r,echo=TRUE}
data$steps[is.na(data$steps)]<-mean(data$steps,na.rm=TRUE)
```
#this is the head of the new data set
```{r head,echo=TRUE}
head(data)
```
#now we create a histogram with the replaced NA values
```{r,echo=TRUE}


sum_steps<-aggregate(data$steps,by=list(data$date),FUN=sum,na.rm=TRUE) 
```

```{r histogram,echo=TRUE,fig.height=4}
hist(sum_steps$x, 
     breaks=seq(from=0, to=25000, by=2500),
     col="green", 
     xlab="Total number of steps", 
     ylim=c(0, 30), 
     main="Total number of steps taken each day\n(NA replaced by mean)")
```
# we now calcluate the mean and median values after replacing the na values
```{r,echo=TRUE}

mean(sum_steps$x)

median(sum_steps$x)

```

#now we create Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

# Convert date into weekdays

```{r,echo=TRUE}
data$days=tolower(weekdays(data$date))
```
#Now categorised days into weekend and weekdays
```{r,echo=TRUE}
data$day_type<-ifelse(data$days=="saturday"|data$days=="sunday","weekend","weekday")
```
#Take mean steps taken on weekend or weekday in the intervals
```{r,echo=TRUE}
avg_steps<-aggregate(data$steps,by=list(data$interval,data$day_type),FUN=mean,na.rm=TRUE)

colnames(avg_steps)<-c("interval","day_type","steps")
```
# Create panel plot between average steps and interval seperated by day type

```{r panelplot, echo=TRUE,fig.height=4}
ggplot(aes(x=interval,y=steps),data=avg_steps)+geom_line()+facet_wrap(~avg_steps$day_type)

```